Theories of deliberated choice

### 2022-10-31 simple.tex
Towards better axioms

An argumentative theory claims that whenever some of its arguments fail to convince $i$ of choosing $\pga$, it can argue something else.
It is said to be truthful iff this claim is correct.
More precisely, its claims are about paths $\hist$ containing ${\arga, \ar_2, \ar_3, …, \ar_k}$ where $\ar_2 \in \dgi(\arga)$, $\ar_3 \in \dgi(\ar_2)$, …, $\ar_k \in \histend$. The following definition makes this precise.
\begin{definition}[Truthfulness]
	\label{def:truth}
	An argumentative theory $(\pga, \arga, \dg)$ for the exposure protocol $(I, \allargs, {\leadsto})$ is \emph{truthful} iff $\forall i \in I, k \in \N^*, (\ar_j)_{j \in \intvl{1, k}} \in \allargs^k, \hist \in \allhist \suchthat [\ar_1 = \arga \land \ar_k \in \histend \land \forall j \in \intvl{1, k}: \ar_j \in R(\hist) \land \forall j \in \intvl{2, k}: \ar_j \in \dgi(\ar_{j - 1}) \land \hist \nileadsto \pga]: [d(\ar_k, \hist) ≠ \emptyset]$.
\end{definition}

>>>>

An argumentative theory claims that whenever some of its arguments fail to convince $i$ of choosing $\pga$, it can argue something else.
It is said to be truthful iff this claim is correct.
The following definition makes this precise.%
\footnote{To be fixed.
If $a_1$, used by the theory, is not decisive, then the theory has an answer, namely, $a_2$, which works at least temporarily.}
\begin{definition}[Truthfulness]
	\label{def:truth}
	An argumentative theory $(\pga, \arga, \dg)$ for the exposure protocol $(I, \allargs, {\leadsto})$ is \emph{truthful} iff $\forall i \in I, \hist \in \allhist, \ar \in \histend, \ar_1 \in \histend: [\hist \nileadsto \pga] ⇒ [\dg(\ar_1, \hist) ≠ \emptyset \land (\hist, \dg(\ar_1, \hist)) \ileadsto \pga]$.
\end{definition}
\begin{axiom}[Answerability]
	\label{ax:ans}
	$\forall \ar, \ar': [\exists \hist \suchthat (\hist, \ar, \ar') \nileadsto \pga \land \exists \hist \suchthat (\hist, \ar, \ar') \ileadsto \pga] ⇒ \exists \hist, \ar_c \suchthat (\hist, \ar, \ar', \ar_c) \nileadsto \pga$.
\end{axiom}
\footnote{$a_1, a_2, a_3$ NOT and $a_1, a_2, a_3, a_4$ YES and $a_1', a_2, a_3, a_4$ NO imply that $a_1, a_2, a_3, a_4, a'_1$ NO.
Generalizing: $\alpha$ NO and $\alpha, a_4$ YES and $\alpha', a_4$ NO imply that $\exists a' \in \alpha' \suchthat \alpha, a_4, a'$ NO.
Let’s try with this too powerful axiom, but then we should weaken it, as it could be that $a_1$ AND $a_4$ TOGETHER are decisive, whence the axiom fails. We should have then some argument $a_1 + a_4$ to replace them, and…?}
\begin{axiom}[Reinstatement]
	\label{ax:r}
	$\forall \hist, \ar_1 \in \gargsi: [(\hist, a_1, a) \nileadsto p \land (\hist, a_1, a, a_2, …) \ileadsto p] ⇒ [(\hist, a_2) \ileadsto p \land \forall \hist': (\hist', a_2, a') \ileadsto p ⇒ (\hist', a_1, a') \nileadsto p]$.
\end{axiom}
\begin{example}[No reinstatement]
	Reinstatement is required. Consider $a_1$ not decisive because of $a$ and then $(a_1, a, a_2) \ileadsto p$. But still $a_2$ not decisive: $a2 \nileadsto p$.
\end{example}
\begin{theorem}[Validity]
	If an argumentative theory $(\pga, \arga, \dg)$ is truthful and $\allargs$ is finite, then \cref{ax:ans,ax:r} imply that the theory is correct.
\end{theorem}
\begin{proof}[Sketch of proof]
	Consider any $i$. We have to show that $\pga \in \Phi_i$.%
	\footnote{Assume $(a_4, \arga, a) \nileadsto \pga$ (thus, $a$ attacks $\arga$). 
	By tf, $(a_4, \arga, a, \ar_2, …) \ileadsto \pga$.
	We get $(a_3, \arga, a, \ar_2) \ileadsto \pga$ by answ.
	By reinst: $(\ar_2)$ YES.}

	Assume that $\arga$ is not decisive (otherwise, done). Thus, for some $\hist: (\hist, \arga, \ar) \nileadsto \pga$ or $(\hist, \ar, \arga)$.
	Consider that $\hist: (\hist, \arga, \ar) \nileadsto \pga$.
	Consider $\hist'$ and wonder whether $(\hist', \arga, \ar) \ileadsto \pga$. If so, 	
	By Answ, this means that $\forall \hist, (\hist, \arga, \ar) \nileadsto \pga$.
	By truthfulness, $\ar_2 = \dg(\ar_1, (\hist, \arga, \ar)) ≠ \emptyset \land (\hist, \arga, \ar, \ar_2) \ileadsto \pga$.
	Assume now that $(\hist, \arga, \ar, \ar_2, …) \ileadsto \pga$.
	Then by Reinstatement, $(\hist, \ar_2) \ileadsto \pga$.

%	Assume, to start with, that the theory is of depth 2, more precisely, $\arga$ is not decisive because $\hist \nileadsto \pga, \arga \in \histend$ and $\dgi(\arga) = \ar_2$ with no further hist against it.
\end{proof}
